Splints of root systems for special Lie subalgebras

### 2014-12-19 AffineLieAlgebras/Splints/special-paper.tex
More on splints

(From wiki:http://en.wikipedia.org/wiki/Minuscule_representation)
Each simple Lie algebra has a unique quasi-minuscule representation that is not minuscule, and the
multiplicity of the zero weight is the number of short nodes of the Dynkin diagram.

@@

What we really need is the classification of all the representations where multiplicities of all
non-zero weights are equal to one. Fortunately, such a classification was obtained in
\cite{plotkin1998visual}. It consists of multiplicity-free representations and adjoint
representations of algebras $B_{r}, C_{r}$ and $F_{4}$.  (? check with
\cite{berenshtein1990multiplicity}).

(See also https://projecteuclid.org/euclid.jmsj/1180135505 for multiplicity-free branching rules)

Having obtained this classification we need to check which weight diagrams can be obtained as the
projections of $\gf$ root system. It is the case for $B_{n}\to D_{n+1}$ or $A_{1}\oplus\dots\oplus
A_{1}\to D_{n+1}$.

Projection of $\gf$ root system to the root space of $\af$ in many cases can be
encoded by augmented Dynkin diagram with simple root multiplicities not greater or equal than 1.
Number of roots in projection is equal to the number of roots in $\gf$ root system (We do
not consider the case of orthogonal roots here). The rank of augmented Dynkin diagram is the same as
the rank of subalgebra $\af$. Having Lie algebra $\af$ and augmented Dynkin
diagram of the same rank it should be possible to reconstruct the embedding $\af\to
\gf$, though not always in a unique way.

We must also account for the case of roots of parallel roots of different length. $BC_{1} $ is the
simplest system with parallel roots, one of its two parallel is twice longer than another. Such a
system appear in study of affine

 The simplest
system with parallel roots have roots of two different length and Such a systems
appear  $BC_{1} $ and its generalization. (We have the dimension of representation of $\af$ and
projection of roots of $\gf$. It's not enough since in general there are different non-equivalent
representations of $\af$. But the number of simple root systems with given number of roots is
finite, so it's possible to check which one has given projection).

Classification of all splints for special embeddings of a given algebra $\af$ is given by
the augmented Dynkin diagrams of the same rank. (!)

Case of diagrams with multiplicities is not particularly interesting since we have just multiple
copies of the same stem (?).

@@

@@

>>>>

(From wiki:\url{http://en.wikipedia.org/wiki/Minuscule_representation})
Each simple Lie algebra has a unique quasi-minuscule representation that is not minuscule, and the
multiplicity of the zero weight is the number of short nodes of the Dynkin diagram.

@@

What we really need is the classification of all the representations where multiplicities of all
non-zero weights are equal to one. Fortunately, such a classification was obtained in
\cite{plotkin1998visual}. It consists of multiplicity-free representations and adjoint
representations of algebras $A_{r}, B_{r}, C_{r}$ and $F_{4}$.  (? check with
\cite{berenshtein1990multiplicity}).

For adjoint representation of algebra $A_{r}$ we see that projection $\Delta'$ of root system
$\Delta_{\gf}$ should consist of roots of subalgebra $\af=A_{r}$ with multiplicity 2. This is the
case for $\gf=A_{r+1}$, so for the special embedding $A_{r}\to A_{r+1} (su(r+1)\to su(r+2))$ we get
a splint.

(See also https://projecteuclid.org/euclid.jmsj/1180135505 for multiplicity-free branching rules)
%% 
%% Having obtained this classification we need to check which weight diagrams can be obtained as the
%% projections of $\gf$ root system. It is the case for $B_{n}\to D_{n+1}$ or $A_{1}\oplus\dots\oplus
%% A_{1}\to D_{n+1}$.
%% 
%% 
%% 
%% Projection of $\gf$ root system to the root space of $\af$ in many cases can be
%% encoded by augmented Dynkin diagram with simple root multiplicities not greater or equal than 1.
%% Number of roots in projection is equal to the number of roots in $\gf$ root system (We do
%% not consider the case of orthogonal roots here). The rank of augmented Dynkin diagram is the same as
%% the rank of subalgebra $\af$. Having Lie algebra $\af$ and augmented Dynkin
%% diagram of the same rank it should be possible to reconstruct the embedding $\af\to
%% \gf$, though not always in a unique way. 
%% 
%% We must also account for the case of roots of parallel roots of different length. $BC_{1} $ is the
%% simplest system with parallel roots, one of its two parallel is twice longer than another. Such a
%% system appear in study of affine 
%% 
%% 
%%  The simplest
%% system with parallel roots have roots of two different length and Such a systems
%% appear  $BC_{1} $ and its generalization. (We have the dimension of representation of $\af$ and
%% projection of roots of $\gf$. It's not enough since in general there are different non-equivalent
%% representations of $\af$. But the number of simple root systems with given number of roots is
%% finite, so it's possible to check which one has given projection).
%% 
%% Classification of all splints for special embeddings of a given algebra $\af$ is given by
%% the augmented Dynkin diagrams of the same rank. (!)
%% 
%% Case of diagrams with multiplicities is not particularly interesting since we have just multiple
%% copies of the same stem (?).
%%

Having obtained the complete classification of splints of projections of root system for special
embeddings, we need to check when such splints lead to the coincidence of branching coefficients
with weight multiplicities in modules of algebra $\sfr$. This depends on the decomposition of
projection of singular element $\pi_{\af}\left(\Psi^{\mu}_{\gf}\right)$ into the combination of
singular elements of subalgebra $\sfr$. Relation between Weyl groups of $\af$ and $\sfr$ is crucial
here.

@@

As the result we see that branching coefficients coincide with weight multiplicities for
the embeddings $something\to something$ and do not coincide for $something else\to something else$.

@@

Computation of branching coefficients is important for different physical models with symmetry
breaking. This computation is drastically simplified if branching coefficients coincide with weight
multiplicities, since efficient Freudental formula can be used \cite{moody1982fast}. Classification
of splints for regular and special embeddings gives us all the cases when this coincidence takes
place. Aside from computational importance, this coincidence is very interesting from
representation-theoretic point of view, because it follows from new unexpected connection between
different simple Lie algebras.
